---
title: "Statistical Inference Part 1 - Course Project"
author: "Clay Burns"
date: "June 18, 2015"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=7, fig.height=4, warning=FALSE, message=FALSE)
```

## Purpose 
This report will explore the exponential distribution in R and compare it with the Cnetral Limit Theorem. This distribution can be simulated in R with rexp(n, lambda) where lambda is the rate paratmeter. The mean of the distrubtion is 1/lamdba and the stand deviation is also 1/lambda. The set lambda = .2 for all of the simulations. The exploration will cover a distribution of 40 exponentials and have a thousand simlultions.

## Simulations

```{r}
# load neccesary libraries
library(ggplot2)

# set constants n(40) = exponetial, lambda(.2) for rexp, number of tests = 1000  
lambda <- 0.2
n <- 40 
numberOfSimulations <- 1000 

# set the seed to create reproducability
set.seed(22678979)

# run the test resulting in n(40) x numberOfSimulations(1000) matrix
exponentialDistributions <- matrix(data=rexp(n * numberOfSimulations, lambda), nrow=numberOfSimulations)
exponentialDistributionMeans <- data.frame(means=apply(exponentialDistributions, 1, mean))
```

```{r echo=FALSE}
# plot the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1) +   
  scale_x_continuous(breaks=round(seq(min(exponentialDistributionMeans$means), max(exponentialDistributionMeans$means), by=1)))
```

## Sample Mean versus Theoretical Mean

The expected mean $\mu$ of a exponential distribution of rate $\lambda$ is 

$\mu= \frac{1}{\lambda}$ 

```{r}
mu <- 1/lambda
mu
```

Let $\bar X$ be the average sample mean of 1000 simulations of 40 randomly sampled exponential distributions.

```{r}
meanOfMeans <- mean(exponentialDistributionMeans$means)
meanOfMeans
```

The two figures are very close. 


## Sample Variance versus Theoretical Variance

The expected standard deviation $\sigma$ of a exponential distribution of rate $\lambda$ is 

$\sigma = \frac{1/\lambda}{\sqrt{n}}$ 

The e

```{r}
sd <- 1/lambda/sqrt(n)
sd
```

The variance $Var$ of standard deviation $\sigma$ is

$Var = \sigma^2$ 

```{r}
Var <- sd^2
Var
```

Let $Var_x$ be the variance of the average sample mean of 1000 simulations of 40 randomly sampled exponential distribution, and $\sigma_x$ the corresponding standard deviation.
```{r}
sd_x <- sd(exponentialDistributionMeans$means)
sd_x

Var_x <- var(exponentialDistributionMeans$means)
Var_x
```

Both figures are quite close but since variance is squared minor difference will likelyy be enhanced more.

## Distribution
Time to compare the population means & standard deviation with a normal distribution of the expected values. Lines will be added for the calculated and expected means

```{r echo=FALSE}
# plot the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1, aes(y=..density..), alpha=0.2) + 
  stat_function(fun = dnorm, arg = list(mean = mu , sd = sd), colour = "red", size=1) + 
  geom_vline(xintercept = mu, size=1, colour="#CC0000") + 
  geom_density(colour="blue", size=1) +
  geom_vline(xintercept = meanOfMeans, size=1, colour="#0000CC") + 
  scale_x_continuous(breaks=seq(mu-3,mu+3,1), limits=c(mu-3,mu+3)) 
```

The graph show just how true the central limit theorm is by how nicely the mean lines line up. 
