---
title: "Statistical Inference Course Project Part 2"
author: "Clay Burns"
date: "June 18, 2015"
output: pdf_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=10, fig.height=8, warning=FALSE, message=FALSE)
```

# Project Overview
* Load ToothGrowth Data & perform basic exploratory data analyses
* Provide basic summary of the data
* Use confidence interval & hypothesis test to compare tooth growth by supp and dose 
* Draw conclusions and list assumptions 
## Load Data & Libraries
```{r}
#Load neccesary libraries
library(ggplot2)
library(datasets)
library(gridExtra)
library(GGally)

# Load Vitamin C on Tooth Growth in Guinea Pigs & convert dose to factor
data(ToothGrowth)
toothGrowth <- ToothGrowth 
toothGrowth$dose <- as.factor(toothGrowth$dose)
```
## Simple Summary of the data
```{r}
str(toothGrowth)
summary(toothGrowth)
head(toothGrowth)

table(toothGrowth$supp, toothGrowth$dose)
```

```{r echo=FALSE}
```{r echo=FALSE}
p1 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~supp) 

p2 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~dose) 

p3 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot()

p4 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot()

grid.arrange(p1, p4, p2, p3, ncol = 2, nrow=2)

```
```{r}
anova.out <- aov(len ~ supp * dose, data= toothGrowth )
summary(anova.out)
```

The results show that there is an interaction between length (len) and dosage (dose) due to a high F value(.000231) and low P value(.000231).

```{r}
TukeyHSD(anova.out)
```
The Tukey analysis show that there are significant differences between each of the group in supp and dose. The only only interaction that lack this fall between (VC:0.5-OJ:0.5; VC:1-OJ:0.5; OJ:2-OJ:1; VC:2-OJ:1 and VC:2-OJ:2).

```{r}
confint(anova.out)
print(model.tables(anova.out,"means"),digits=3)
```
## Conclusions
Both supplement and dosage have clear indipendent effects on the length of teeth for guinea pigs. Holding things constant it means longer teeth on average. The supplement type has a clear influence  but OJ has a greater average teethgrowth in combination with dosage .5 and 1 then for the VC supplement. While at higher levels there is no significant effect. 

These assumpitions still have to be dealt with. 

* That the guinea pigs are attempting to be repesentative for the whole population
* That dosage and supplement were properly assigned
* That the distribution of the means is normal. 